Stack Safety as a Security Property

### 2020-07-04 main.tex
From call
done with intro for now; releasing lock

%% Title information
\title{Stack Safety as a Security Property}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'

@@

The call stack is an increasingly\apt{``a perennially'' ?} popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from attackers reading or corrupting private stack data, to
highjacking a program's control flow. To prevent such attacks, many
techniques have been proposed in the literature\leo{cite, point out?
  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
  stronger point.} aiming to achieve ``stack
  safety''.

But what {\em is} this stack safety of which we speak? A precise formal
definition is lacking in the literature, with the notion usually being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other
  approaches}\bcp{Yes, we have to be very careful about this}.
\leo{Examples at a high level}
Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
  we are only aware of one, quite recent, attempt to do something better} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
but even here, stack safety is defined with respect to an abstract
machine that ``satisfies stack safety by construction,'' begging
the question of what {\em is} the property that holds by construction?
\apt{Suggest to move first sentence of next para up to this one.}

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. The first
contribution of the present paper is to
turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
  claim this} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully
developed.}\bcp{Yes, and let's remember that they are highly likely to be
reviewing this paper!  Let's be as generous as we can be without
under-selling our delta.}

The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property with both confidentiality and
integrity components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}

\bcp{``alternative framework that uses the conceptual framework of language-based security''}
\bcp{idea:intrinsic vs. extrinsic: stktokens only talks}

Another dimension: spatial vs temporal safety

While integrity and confidentiality are relatively standard security
properties, the stack-safety setting raises an interesting additional
challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be proteced from {\em its} callees.
We address this challenge by varying the access powers
of the attacker at call points.

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety between
functions: all code is uniquely owned by some function, and all jumps between functions
are either calls that enter at a marked entry point or returns that enter
just after their associated call in the standard nested call structure.

 and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
  the sentence are about two very different things.  The second deserves its
  own paragraph, IMO.}\apt{+1}
%\leo{There needs to be some informal description
%  of the wbcf part here. Sean?}

In summary, we offer the following contributions\apt{is this up to date with current paper organization?}:
\begin{itemize}
\item \leo{More? Extensional}\bcp{We might need to invent a word for
  this---''quasi-extensional''?} formal characterizations of the
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\Cref{sec:lse-and-wbcf}). \leo{Make one sentence} examples from the literature also fail our
  property.\apt{Something about relation to StkTokens formalization?}
\item Case Study:
  \leo{Existing enforcement (Nick-eager). Satisfies Property.
  Satisfies even stronger property (provably).}
\item Pragmatically motivated refinement:
  \leo{Lazy enforcement mechanism (Nick-lazy).
    Plus a lazy variant of the property.}
%\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with a running example that we're going to keep coming back
to throughout the paper (Section~\ref{sec:running-example}) and some
technical preliminaries (Section~\ref{sec:prelim}).

@@

%% Bibliography
%\bibliography{bcp.bib}

>>>>

%% Title information
\title{Stack Safety as a Security Property}         %% [Short Title] is optional;
                                        %% when present, will be used in
v                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'

@@

The call stack is a perennially popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
hijacking of the program's control flow. To prevent such attacks, many
software and hardware techniques for stack safety have been proposed, % in the literature and implemented in production systems,
including stack canaries~\citep{Cowan+98},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
split stacks~\citep{Kuznetsov+14},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from nick;there could be more}

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion of safety
is usually defined negatively, by enumerating the bad things that a safe system
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits 
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ... 
We are only aware of one, quite recent, attempt to frame a positive definition.
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, local state encapsulation and well-bracketed control flow.
They formalize these properties by defining an abstract capability-based machine   
that ``satisfies [stack safety] by construction.'' We believe there is still room
to craft a more abstract and less intensional characterization of
stack safety, and that is the goal of the present paper. \apt{still needs tuning}

Our approach is to give alternative formal
definitions of local state encapsulation and well-bracketed control flow
using the conceptual framework of language-based security.
%
The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property with both confidentiality and
integrity components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}

% BCP idea: contrast intrinsic vs. extrinsic property definition
% Remember to talk about spatial vs temporal safety

While integrity and confidentiality are standard security
properties, applying them in the stack-safety setting raises a novel challenge:
the notion of ``attacker'' changes during execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be protected from {\em its} callees.
We address this challenge by adjusting the integrity and confidentiality
of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}

The underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety between
functions: each instruction is uniquely owned by some function, and all jumps between functions
are either calls that enter at a marked entry point or returns that re-enter
just after their associated call in the standard nested call structure.
\apt{clarify what we are contributing here.}

To provide evidence that our definitions capture the inituitive notion of stack safety,
we show that they rule out a wide \apt{really?} range of exploitable bad behaviors.\apt{``from the literature''??}
To assess whether they are implementable, 
we compare them against the proposed capability-based enforcement
mechanism of \citet{Skorstengaard+19} \apt{we really should} and the existing tag-based enforcement mechanism
of \citet{DBLP:conf/sp/RoesslerD18}.

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the 
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\Cref{sec:lse-and-wbcf}) in a very general setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\Cref{sec:enforcement}).
  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing property?}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy 
  of \citet{DBLP:conf/sp/RoesslerD18}~(\Cref{sec:lazy}).  This policy does not enforce our property, but we give a
  variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer
to throughout the paper (Section~\ref{sec:running-example}) and some
technical preliminaries (Section~\ref{sec:prelim}).

@@

%% Bibliography
\bibliography{bcp.bib,local.bib}
